---
title: "EDA"
format:
  html:
    code-fold: true
---

## Penguin Size and Mass by Sex and Species


```{r}
library(palmerpenguins)
library(dplyr)
library(ggplot2)

df <- palmerpenguins::penguins
```


```{r}
df %>%
  group_by(species, sex) %>%
  summarise(
    across(
      where(is.numeric), 
      \(x) mean(x, na.rm = TRUE)
      )
    ) %>%
  knitr::kable()
```

## Penguin Size vs Mass by Species
This graph compares the average body mass of the penguins to their bill lengths. It can be seen that the species have pretty clear groups - for example, gentoo penguins seem to have higher body masses and bill lengths than adelle penguins.

```{r}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

## Penguin Species Model with LDA
This model seeks to predict penguin species by their bill length.

```{r}
library(tidymodels) 
penguin_split <- initial_split(df, prop = 0.8)
penguin_train <- training(penguin_split)
penguin_test <- testing(penguin_split)
lda_spec <- discrim_linear() %>%
        set_mode("classification") %>%
        set_engine("MASS")
lda_fit<- lda_spec %>%
    fit(species~bill_length_mm, data = penguin_train)
lda_fit_preds_test <- augment(lda_fit , new_data = penguin_test) %>%
        select(species, bill_length_mm, .pred_class) 
lda_fit_preds_test %>%
  conf_mat(truth = species, estimate = .pred_class)
lr_metrics <- metric_set(accuracy, sensitivity, specificity, ppv)
lda_fit_preds_test%>%
  lr_metrics(truth = species, estimate = .pred_class, event_level = "second")
```
We can see here that the model is not the best at predicting the penguins' species. It has decent specificity, but not much else.